---
title: "Graph_Joyce"
author: "Qiaoyi Xu"
date: "2022-11-27"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r, load data}
info_scatter = read_csv("data/nsduh_19.csv")

info_scatter
```



```{r, scatter plot}
info_scatterplot = info_scatter %>%
  filter(irmjage!=991)%>%
  group_by(irmjage) %>%
  summarise(number_people = n()) %>%
  ggplot(aes (x = irmjage, y = number_people)) +
  geom_point() +
  geom_smooth() +
  labs(x = "Age of first used marijuana", y = "Frequency(number of people)",
       title = "Age When First Used among people who used MARIJUANA/HASHISH")
  

info_scatterplot                                 
      
```

In this part, we created a scatter plot to show the age when first used MARIJUANA/HASHISH and frequency of people in each year. We excluded people who never use MARIJUANA/HASHISH in this plot. Because, without doubts, most people never use MARIJUANA/HASHISH. In our dataset, there are 31340 people who never use MARIJUANA/HASHISH. Therefore, after excluded poeple who never use MARIJUANA/HASHISH, it is more clear to see the distribution of Age when First Used among people who used MARIJUANA/HASHISH.     


From this plot, we could conclude that most people who used Marijuan strarted bwtween 10 and 30 years old when first used marijuana. So, it is necessary to focus on youth population during our further steps. 





















```{r}
info_scatterplot_addneveruse = info_scatter %>%
  mutate(irmjage = as.character(irmjage),
         irmjage = recode(irmjage, "991" = "0"), #0 = "Never Use"
         irmjage = as.numeric(irmjage))%>%
  group_by(irmjage) %>%
  summarise(number_people = n()) %>%
  ggplot(aes (x = irmjage, y = number_people)) +
  geom_point() +
  theme_grey() +
  labs(x = "Age of first used marijuana", y = "Frequency(number of people)",
       title = "Age When First Used MARIJUANA/HASHISH vs frequency(included never use)")
  

info_scatterplot_addneveruse  
```













=============================================================================================================================  
not useful code, prepare for future use if need:

  mutate(firstuse_age = if_else(irmjage<12, "Under 12",
                                   if_else(irmjage>=12&irmjage<=18, "12-18",
                                   if_else(irmjage>=19&irmjage<=25, "19-25", 
                                   if_else(irmjage>=26&irmjage<=30, "26-30",
                                   if_else(irmjage>=31&irmjage<=40, "31-40",
                                   if_else(irmjage>=41&irmjage<=50, "41-50",
                                   if_else(irmjage>=51&irmjage<=60, "51-60", 
                                   if_else(irmjage>60&irmjage<=82, "above 60", 
                                   if_else(irmjage == 991, "Never Use", NA_character_))))))))))
                                   






