---
title: "Mapping"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE}
library(tidyverse)
library(dplyr)
library(usmap)
library(ggplot2)
require(maps)
library(plotly)

```

### Where marijuana is legal in the United States?

```{r, warning = FALSE}
legal_state = read_csv("data/legalization_state.csv") 

legal_state = legal_state %>% 
  rename(full = State)

legal_state = legal_state %>%
  mutate(status = case_when(
    Recreational == "Yes" & Medical == "Yes" ~ "Recreational/Medical", #rec/med
    Recreational == "No" & Medical == "Yes" ~ "Medical", #medical legal
    Recreational == "No" & Medical == "No" ~ "None"#none
  )) 

```



```{r, warning = FALSE}
data(statepop)

mapping_df <-inner_join(statepop, legal_state, by = "full")

mapping_df <- mapping_df %>%
  select(-pop_2015) %>%
  mutate(status = factor(status))

labels = usmapdata::centroid_labels("states")
state_labels <- merge(mapping_df, labels, by = "fips")

map = plot_usmap(regions = "states", data=mapping_df, values = "status",color = "white", labels = TRUE)+
  theme(panel.background = element_rect(colour = "black")) +
  scale_fill_manual(values = c(`Recreational/Medical` = "red", `Medical` = "blue", `None` = "grey" ), name = "Legalized states") + 
  theme(legend.position = "bottom") +
  labs(title = "Where marijuana is legal in the United States")

map$layers[[2]]$aes_params$size <- 3
```

```{r, echo=FALSE,eval=FALSE}
plot_usmap(regions = "states", data=mapping_df, values = "status")+ 
  ggrepel::geom_label_repel(data = labels,
             aes(x = x, y = y, label = abbr),
             size = 3, alpha = 0.8,
             label.r = unit(0.5, "lines"), label.size = 0.5,
             segment.color = "red", segment.size = 1,
             seed = 1002) +
  theme(panel.background = element_rect(colour = "black")) +
  scale_fill_manual(values = c(`Recreational/Medical` = "red", `Medical` = "blue", `None` = "grey" ), name = "Legalized states") + 
  theme(legend.position = "bottom", ) +
  labs(title = "Where marijuana is legal in the United States")
```


```{r, echo=FALSE, eval=FALSE}
library(ggiraph)

ggplot(mapping_df) +
  geom_sf_interactive(
    aes(fill = status,
        tooltip = sprintf("%s<br/>%s", Year_legalized_REC, Year_legalized_MED))) +
  theme_wsj(color = "white", base_family = "mono") + 
  ggtitle("Places I've been to") +
  scale_fill_manual(breaks = c("Recreational/Medical", "Medical", "None"), 
                      labels=c("Recreational/Medical", "Medical", "None")) +
  guides(fill=guide_legend(title=NULL))

x <- girafe(ggobj = gg)
x <- girafe_options(x,
                    opts_zoom(min = 0.7, max = 2) )
if( interactive() ) print(x)
```


```{r, warning = FALSE}
ggplotly(map) %>%
  highlight("plotly_hover", selected = attrs_selected(line = list(color = "black"))) 
```

### How many states legalized for marijuana use?
```{r, warning = FALSE}
legalized_states <- 
  legal_state%>% 
  select(-Recreational, -Medical) %>% 
  mutate(status = fct_relevel(status, "Recreational/Medical", "Medical", "None")) %>%
  group_by(status) %>%
  summarise(state_number = n()) %>%
  knitr::kable(col.names = c("State", "number of states"))

legalized_states
```

### How many people could use marijuana legally in 2019?




### When marijuana legaled in each State?

```{r, warning = FALSE}
year_df = mapping_df %>%
  select(full,Year_legalized_REC,Year_legalized_MED) %>%
  pivot_longer(
    Year_legalized_REC:Year_legalized_MED,
    names_to = "rec_med", 
    values_to = "year")


legal_year <- year_df %>%
  plot_ly(x = ~ year, y = ~full, color = ~rec_med, type = "scatter", mode = "markers") %>%
  layout(title = "Legalized year of each state", 
         yaxis = list(title = "States", tickfont = list(size = 5)), xaxis = list(title = "Legalized year")) #add title in the graph

legal_year

```


library(plotly)
mapping_df %>% plot_ly(
    type = 'scatter',
    mode = 'markers',
    x = ~Year_legalized_REC,
    y = ~Year_legalized_MED,
    color = ~status,
    text = ~full,
    hovertemplate = paste(
      "<b>%{text}</b><br><br>",
      "%{yaxis.title.text}: %{y:$,.0f}<br>",
      "%{xaxis.title.text}: %{x:.0%}<br>",
      "Number Employed: %{marker.size:,}",
      "<extra></extra>"
      ))


    
fig
```
 df <- read.csv('https://raw.githubusercontent.com/plotly/datasets/master/job-automation-probability.csv')
 
 https://plotly.com/r/hover-text-and-formatting/ 
 
 
 

add population into the map:
https://cran.r-project.org/web/packages/usmap/vignettes/advanced-mapping.html

https://bhaskarvk.github.io/user2017.geodataviz/notebooks/03-Interactive-Maps.nb.html 


```{r, echo=FALSE, eval=FALSE}

basic_map = ggplot() + 
  geom_polygon( data=MainStates, aes(x=long, y=lat, group=group),
                color="black", fill="lightblue" )

basic_map
```


mapping_df <-inner_join(MainStates, legal_state, by = "region") #更改每个州名字的格式


