---
title: "Statistical Analysis"
output: 
  html_document:
    toc: true
    toc_float: true
---
```{r}
library(tidyverse)
```

# Tidy
```{r}
data = read_csv("./data/nsduh_19.csv")
reg = data %>% 
  mutate(
    ymdelt = replace_na(ymdelt, 0),
         amdelt = replace_na(amdelt, 0),
         drp = ymdelt + amdelt,
         mrjmon = as.factor(mrjmon),
         mrjmon = fct_relevel(mrjmon, "0"),
         drp = na_if(drp, 0),
         drp = as.factor(drp),
         drp = fct_relevel(drp, "2"),
         irsex = as.factor(irsex),
         newrace2 = as.factor(newrace2),
         eduhighcat = as.factor(eduhighcat),
         coutyp4 = as.factor(coutyp4),
         income = as.factor(income),
         irmarit = as.factor(irmarit)
         )
```
drp 0 1 (yes) 2 (no) now two levels
MRJMON 0 No 1 Yes

# Logistic regression
```{r}
reg %>% 
  glm(drp ~ mrjmon + catag6 + irsex + newrace2 + eduhighcat + coutyp4 + income + irmarit, data = ., family = binomial()) %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate),
         conf.low = exp((estimate - 1.96 * std.error)),
         conf.high = exp((estimate + 1.96 * std.error))
         ) %>%
  filter(term == "mrjmon1") %>% 
  select(term, OR, conf.low, conf.high) %>% 
  knitr::kable(digits = 3)
```
The odds of depression is 1.901 (95% CI: 1.794 - 2.015) comparing people who used cannabis in the past month to those who do not adjusting for age, sex, race, education, region, income, and martial status. 

I am not sure which outcome I regess on, need to check using SAS. 

Can we use covairates to predict drp? Maybe do a Cross-Validation 

hypothesise test?

Region (coutyp4) distribution of cannabis smoking and drp





