---
title: "Statistical Analysis"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---
```{r, message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
library(tidyverse)

```

```{r, message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
data = read_csv("./data/nsduh_19.csv")
reg = data %>% 
  mutate(amdeyr = as.factor(amdeyr),
         amdeyr = fct_relevel(amdeyr, "2"),
         ymdeyr = as.factor(ymdeyr),
         ymdeyr = fct_relevel(ymdeyr, "2"),
         mrjmon = as.factor(mrjmon),
         mrjmon = fct_relevel(mrjmon, "0"),
         irsex = as.factor(irsex),
         newrace2 = as.factor(newrace2),
         eduhighcat = as.factor(eduhighcat),
         coutyp4 = as.factor(coutyp4),
         income = as.factor(income),
         irmarit = as.factor(irmarit)
         )
```

## Depression by sociodemographic characteristcs
### Youths (12 - 17 years old)
1. Youth depression by age category: 
```{r}
data %>% 
  select(catag6, ymdeyr) %>% 
  drop_na(ymdeyr) %>% 
  mutate(ymdeyr = ifelse(ymdeyr == 2, 0, 1),
         catag6 = recode(catag6, "1" = "12-17 years old",
                         "2" = "18-25 years old",
                         "3" = "26-34 years old",
                         "4" = "35-49 years old",
                         "5" = "50-64 years old",
                         "6" = "65+ years old"
                         )) %>% 
  group_by(catag6) %>% 
  summarise(
    prevalence_of_past_year_depression = sum(ymdeyr, na.rm = TRUE) / n() * 100) %>% 
  knitr::kable(digits = 2)
```
* The prevalence of past year major depressive episode for youth is 16.2%.   
 

2. Youth depression by gender: 
```{r}
data %>% 
  select(irsex, ymdeyr) %>% 
  drop_na(ymdeyr) %>% 
  mutate(ymdeyr = ifelse(ymdeyr == 2, 0, 1),
         irsex = recode(irsex, "1" = "Male",
                         "2" = "Female")) %>% 
  group_by(irsex) %>% 
  summarise(
    prevalence_of_past_year_depression = sum(ymdeyr, na.rm = TRUE) / n() * 100) %>% 
  knitr::kable(digits = 2)
```
* The prevalence of past year major depressive episode for youth is highest among female.
* Female youth depression is more than twofold of male youth depression.

3. Youth depression by race:
```{r}
data %>% 
  select(newrace2, ymdeyr) %>% 
  drop_na(ymdeyr) %>% 
  mutate(ymdeyr = ifelse(ymdeyr == 2, 0, 1),
         newrace2 = recode(newrace2, "1" = "NonHisp White",
                         "2" = "NonHisp Black",
                         "3" = "NonHisp Native Am",
                         "4" = "NonHisp Native HI",
                         "5" = "NonHisp Asian",
                         "6" = "NonHisp more than one race",
                         "7" = "Hispanic"
                         )) %>% 
  group_by(newrace2) %>% 
  summarise(
    prevalence_of_past_year_depression = sum(ymdeyr, na.rm = TRUE) / n() * 100) %>% 
  knitr::kable(digits = 2)
```
* The prevalence of past year major depressive episode for youth is highest among those who identified them to have more than one race. 
* The lowest is among Non-Hispanic Black. 


4. Youth depression by family income:  
```{r}
data %>% 
  select(income, ymdeyr) %>% 
  drop_na(ymdeyr) %>% 
  mutate(ymdeyr = ifelse(ymdeyr == 2, 0, 1),
         income = recode(income, "1" = "Less than $20,000",
                         "2" = "$20,000 - $49,999",
                         "3" = "$50,000 - $74,999",
                         "4" = "$75,000 or More"
                         )) %>% 
  group_by(income) %>% 
  summarise(
    prevalence_of_past_year_depression = sum(ymdeyr, na.rm = TRUE) / n() * 100) %>% 
  knitr::kable(digits = 2)
```
* The prevalence of past year major depressive episode for youth is highest among those with total family income range $50,000 - $74,999.


### Adults (18 + years old)
1. Adult depression by age categories: 
```{r}
data %>% 
  select(catag6, amdeyr) %>% 
  drop_na(amdeyr) %>% 
  mutate(amdeyr = ifelse(amdeyr == 2, 0, 1),
         catag6 = recode(catag6, "1" = "12-17 years old",
                         "2" = "18-25 years old",
                         "3" = "26-34 years old",
                         "4" = "35-49 years old",
                         "5" = "50-64 years old",
                         "6" = "65+ years old"
                         )) %>% 
  group_by(catag6) %>% 
  summarise(
    prevalence_of_past_year_depression = sum(amdeyr, na.rm = TRUE) / n() * 100) %>% 
  knitr::kable(digits = 2)
```
* The highest prevalence of past year major depressive episode for adult is among those aged 18 - 25 years old.  
* And the prevalence decreases with age.
  
 

2. Adult depression by gender:
```{r}
data %>% 
  select(irsex, amdeyr) %>% 
  drop_na(amdeyr) %>% 
  mutate(amdeyr = ifelse(amdeyr == 2, 0, 1),
         irsex = recode(irsex, "1" = "Male",
                         "2" = "Female")) %>% 
  group_by(irsex) %>% 
  summarise(
    prevalence_of_past_year_depression = sum(amdeyr, na.rm = TRUE) / n() * 100) %>% 
  knitr::kable(digits = 2)
```
* The prevalence of past year major depressive episode for adult is also highest among female. However, the difference is not as large as that of the youth. 
  


3. Adult depression by race:
```{r}
data %>% 
  select(newrace2, amdeyr) %>% 
  drop_na(amdeyr) %>% 
  mutate(amdeyr = ifelse(amdeyr == 2, 0, 1),
         newrace2 = recode(newrace2, "1" = "NonHisp White",
                         "2" = "NonHisp Black",
                         "3" = "NonHisp Native Am",
                         "4" = "NonHisp Native HI",
                         "5" = "NonHisp Asian",
                         "6" = "NonHisp more than one race",
                         "7" = "Hispanic"
                         )) %>% 
  group_by(newrace2) %>% 
  summarise(
    prevalence_of_past_year_depression = sum(amdeyr, na.rm = TRUE) / n() * 100) %>% 
  knitr::kable(digits = 2)
```
* The prevalence of past year major depressive episode for adult is highest among those who identified them to have more than one race. 
* The lowest is among Non-Hispanic Native Hawiian. 
 

4. Adult depression by education:
(We are not considering youth education in this case.)
```{r}
data %>% 
  select(eduhighcat, amdeyr) %>% 
  drop_na(amdeyr) %>% 
  mutate(amdeyr = ifelse(amdeyr == 2, 0, 1),
         eduhighcat = recode(eduhighcat, "1" = "Less high school",
                         "2" = "High school grad",
                         "3" = "Some coll/Assoc",
                         "4" = "College graduate"
                         )) %>% 
  group_by(eduhighcat) %>% 
  summarise(
    prevalence_of_past_year_depression = sum(amdeyr, na.rm = TRUE) / n() * 100) %>% 
  knitr::kable(digits = 2)
```
* The prevalence of past year major depressive episode for adult is highest among those with some college degrees.
 


5. Adult depression by family income:  
```{r}
data %>% 
  select(income, amdeyr) %>% 
  drop_na(amdeyr) %>% 
  mutate(amdeyr = ifelse(amdeyr == 2, 0, 1),
         income = recode(income, "1" = "Less than $20,000",
                         "2" = "$20,000 - $49,999",
                         "3" = "$50,000 - $74,999",
                         "4" = "$75,000 or More"
                         )) %>% 
  group_by(income) %>% 
  summarise(
    prevalence_of_past_year_depression = sum(amdeyr, na.rm = TRUE) / n() * 100) %>% 
  knitr::kable(digits = 2)
```
* The prevalence of past year major depressive episode for adult is highest among those with total family income range less than $20,000. 
 

6. Adult depression by marital status:  
```{r}
data %>% 
  select(irmarit, amdeyr) %>% 
  drop_na(amdeyr) %>% 
  mutate(amdeyr = ifelse(amdeyr == 2, 0, 1),
         irmarit = recode(irmarit, "1" = "Married",
                         "2" = "Widowed",
                         "3" = "Divorced or Separated",
                         "4" = "Never Been Married"
                         )) %>% 
  group_by(irmarit) %>% 
  summarise(
    prevalence_of_past_year_depression = sum(amdeyr, na.rm = TRUE) / n() * 100) %>% 
  knitr::kable(digits = 2)
```
* The prevalence of past year major depressive episode for adult is highest among who have never been married. 
* And is lowest among those who is married. 
 

## Logistic regression and Hypothesis testing
### Youths (12 - 17 years old)  
1. We perform a logistic regression on our outcome and exposure on youths while adjusting for other covariates.  
ymdeyr (youth depression) = mrjmon (Past month marijuana Use) + catag6 (age) + irsex (sex) + newrace2 (race) + coutyp4 (region)  
And the resulting odds ratio is shown below:
```{r}
reg %>% 
  glm(ymdeyr ~ mrjmon + catag6 + irsex + newrace2 + coutyp4, data = ., family = binomial()) %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate),
         conf.low = exp((estimate - 1.96 * std.error)),
         conf.high = exp((estimate + 1.96 * std.error))
         ) %>%
  filter(term == "mrjmon1") %>% 
  select(term, OR, conf.low, conf.high) %>% 
  knitr::kable(digits = 3)
```
* The odds of depression is 2.673 (95% CI: 2.309 - 3.093) comparing youth who used cannabis in the past month to those who do not adjusting for age, sex, race, region. 
 

2. Is the beta for our exposure variable truly different from zero?  
We obtained the p.value from the t.test statistics.
```{r}
reg %>% 
  glm(ymdeyr ~ mrjmon + catag6 + irsex + newrace2 + coutyp4, data = ., family = binomial()) %>% 
  broom::tidy() %>% 
  filter(term == "mrjmon1") %>% 
  select(term, statistic, p.value) %>% 
  knitr::kable(digits = 4)
```
* As we can see that the t-tests statistic for the term `mrjmon1` which is marijuana past month use comparing Yes to No is very large and the resulting p.value is close to zero. Our beta coefficient of the exposure variable is significant. 
 


### Adults (18 + years old)
1. We perform a logistic regression on our outcome and exposure on adults while adjusting for other covariates.  
amdeyr (adult depression) = mrjmon (Past month marijuana Use) + catag6 (age) + irsex (sex) + newrace2 (race)  + eduhighcat (education) + coutyp4 (region) + income (income) + irmarit (martial status)  
And the resulting odds ratio is shown below:
```{r}
reg %>% 
  glm(amdeyr ~ mrjmon + catag6 + irsex + newrace2 + eduhighcat + coutyp4 + income + irmarit, data = ., family = binomial()) %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate),
         conf.low = exp((estimate - 1.96 * std.error)),
         conf.high = exp((estimate + 1.96 * std.error))
         ) %>%
  filter(term == "mrjmon1") %>% 
  select(term, OR, conf.low, conf.high) %>% 
  knitr::kable(digits = 3)
```
* The odds of depression is 1.935 (95% CI: 1.795 - 2.086) comparing adults who used cannabis in the past month to those who do not adjusting for age, sex, race, education, region, income, and martial status. 


2. Is the beta for our exposure variable truly different from zero?  
We obtained the p.value from the t.test statistics.
```{r}
reg %>% 
  glm(amdeyr ~ mrjmon + catag6 + irsex + newrace2 + eduhighcat + coutyp4 + income + irmarit, data = ., family = binomial()) %>% 
  broom::tidy() %>% 
  filter(term == "mrjmon1") %>% 
  select(term, statistic, p.value) %>% 
  knitr::kable(digits = 4)
```
* As we can see that the t-tests statistic for the term `mrjmon1` which is marijuana past month use comparing Yes to No is very large and the resulting p.value is close to zero. Our beta coefficient of the exposure variable is significant. 

## Finding & Summary
* Depression was distributed unevenly across a wide range of sociodemographic variables.
* The prevalence of youth depression was higher among female. It is also true for adults. 
* Depression seems to be decreasing with age. The highest prevalence is among those who is youngest.  
* Depression is more prevalent for adults with family income less than $20,000 annually and who have never been married. 
* People who identified them to have more than one race have the highest prevalence of depression among all the race categories for both youth and adults.   
* Depression was significantly more common among those who reported marijuana use in the past month for both youths and adults, adjusting for demographic factors. 
* The odds of depression is 2.673 (95% CI: 2.309 - 3.093) comparing youth who used cannabis in the past month to those who do not adjusting for age, sex, race, region. 
* The odds of depression is 1.935 (95% CI: 1.795 - 2.086) comparing adults who used cannabis in the past month to those who do not adjusting for age, sex, race, education, region, income, and martial status.




















