---
title: "Statistical Analysis"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---
```{r, message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
library(tidyverse)

```

```{r, message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
data = read_csv("./data/nsduh_19.csv")
reg = data %>% 
  mutate(amdeyr = as.factor(amdeyr),
         amdeyr = fct_relevel(amdeyr, "2"),
         ymdeyr = as.factor(ymdeyr),
         ymdeyr = fct_relevel(ymdeyr, "2"),
         mrjmon = as.factor(mrjmon),
         mrjmon = fct_relevel(mrjmon, "0"),
         irsex = as.factor(irsex),
         newrace2 = as.factor(newrace2),
         eduhighcat = as.factor(eduhighcat),
         coutyp4 = as.factor(coutyp4),
         income = as.factor(income),
         irmarit = as.factor(irmarit)
         )
```

# Logistic regression --- Youths (12 - 17 years old)
We perform a logistic regression on our outcome and exposure on youths while adjusting for other covariates.  
ymdeyr (youth depression) = mrjmon (Past month marijuana Use) + catag6 (age) + irsex (sex) + newrace2 (race) + coutyp4 (region)  
And the resulting odds ratio is shown below:
```{r}
reg %>% 
  glm(ymdeyr ~ mrjmon + catag6 + irsex + newrace2 + coutyp4, data = ., family = binomial()) %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate),
         conf.low = exp((estimate - 1.96 * std.error)),
         conf.high = exp((estimate + 1.96 * std.error))
         ) %>%
  filter(term == "mrjmon1") %>% 
  select(term, OR, conf.low, conf.high) %>% 
  knitr::kable(digits = 3)
```
The odds of depression is 2.673 (95% CI: 2.309 - 3.093) comparing youth who used cannabis in the past month to those who do not adjusting for age, sex, race, region. 


# Hypothesis testing --- Youths (12 - 17 years old)
Is the beta for our exposure variable truly different from zero?  
We obtained the p.value from the t.test statistics.
```{r}
reg %>% 
  glm(ymdeyr ~ mrjmon + catag6 + irsex + newrace2 + coutyp4, data = ., family = binomial()) %>% 
  broom::tidy() %>% 
  filter(term == "mrjmon1") %>% 
  select(term, statistic, p.value) %>% 
  knitr::kable(digits = 4)
```
As we can see that the t-tests statistic for the term `mrjmon1` which is marijuana past month use comparing Yes to No is very large and the resulting p.value is close to zero. Our beta coefficient of the exposure variable is significant. 


# Logistic regression --- Adults (18 + years old)
We perform a logistic regression on our outcome and exposure on adults while adjusting for other covariates.  
amdeyr (adult depression) = mrjmon (Past month marijuana Use) + catag6 (age) + irsex (sex) + newrace2 (race)  + eduhighcat (education) + coutyp4 (region) + income (income) + irmarit (martial status)  
And the resulting odds ratio is shown below:
```{r}
reg %>% 
  glm(amdeyr ~ mrjmon + catag6 + irsex + newrace2 + eduhighcat + coutyp4 + income + irmarit, data = ., family = binomial()) %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate),
         conf.low = exp((estimate - 1.96 * std.error)),
         conf.high = exp((estimate + 1.96 * std.error))
         ) %>%
  filter(term == "mrjmon1") %>% 
  select(term, OR, conf.low, conf.high) %>% 
  knitr::kable(digits = 3)
```
The odds of depression is 1.935 (95% CI: 1.795 - 2.086) comparing adults who used cannabis in the past month to those who do not adjusting for age, sex, race, education, region, income, and martial status. 


# Hypothesis testing --- Adults (18 + years old)
Is the beta for our exposure variable truly different from zero?  
We obtained the p.value from the t.test statistics.
```{r}
reg %>% 
  glm(amdeyr ~ mrjmon + catag6 + irsex + newrace2 + eduhighcat + coutyp4 + income + irmarit, data = ., family = binomial()) %>% 
  broom::tidy() %>% 
  filter(term == "mrjmon1") %>% 
  select(term, statistic, p.value) %>% 
  knitr::kable(digits = 4)
```
As we can see that the t-tests statistic for the term `mrjmon1` which is marijuana past month use comparing Yes to No is very large and the resulting p.value is close to zero. Our beta coefficient of the exposure variable is significant. 















reg %>% 
mutate(coutyp4 = as.character(coutyp4),
    coutyp4 = recode(coutyp4, 
                     "1" = "Large Metro", "2" = "Small Metro", "3" = "Nonmetro"),
    coutyp4 = fct_relevel(coutyp4, "Large Metro", "Small Metro", "Nonmetro")) %>% 
  group_by(coutyp4) %>% 
    summarise( youth = round(sum(youth_dpr, na.rm=TRUE)/n()*100, 2),
               adult = round(sum(adult_dpr, na.rm=TRUE)/n()*100, 2))




















